Poisson approximate likelihood compared to the particle filter

### 2024-09-17 ms.Rnw

\begin{table}[ht] % The placement specifier can be [h!tbp]
\centering 
\caption{AIC for the OvOv rotavirus model, computed using two filtering methods. PAL is the Poisson approximate likelihood, implemented using the code of WWR. PF is the particle filter, implemented using the R package pomp \citep{pomppackagepaper}. Lines 1, 2 and 7 are from WWR, and the remainder are our own computations. We used $5 \times 10^4$ particle for both PF and PAL. PF was repeated 36 times to reduce the Monte Carlo variance, but this step was not necessary for PAL due to its lower Monte~Carlo variance. PF was maximized using iterated filtering, following the approach of SBH. PAL was maximized using coordinate gradient descent, using the code of WWR.
} 
\label{tab:ovovrealdata}
\begin{tabular}{lllllr} 
  \hline
  & Method & Data & Model & Parameters                          & AIC  \\
  \hline
  1. & PF  & SBH  & SBH   & Table~1 of WWR, originally from SBH &  \Sexpr{myround(20134.38,tab_signif)}
  \\
  2. & PAL & WWR  & WWR   & Table~1 of WWR                      & \Sexpr{myround(13778.08,tab_signif)}
  \\
  3. & PAL & WWR  & WWR   & Re-evaluating the data, parameters from WWR &
    \Sexpr{myround(-2*palfilter_lik_list_wwr$log_lik + 2*7 ,tab_signif)} 
  \\
  4. & PF  & WWR & WWR    & Parameters matching 3. &
    \Sexpr{myround(-2*logmeanexp(logLik(pfilter_wwr)) + 2*7,tab_signif)}
  \\
  5. & PF  & WWR & WWR modified   & Initial values changed to match SBH &
     \Sexpr{myround(-2*est_mle_vec_sbh["logLik"] + 2*7,tab_signif)}
  \\
  6. & PAL & WWR & WWR modified   & Data, model and parameters matching 5.  &
     \Sexpr{myround(-2*palfilter_mif2_mle$log_lik + 2*7 ,tab_signif)}
  \\   
  7. & ARMA & SBH & ARMA benchmark    & Sec.~7.3.2 of WWR & 23043
  \\
  8. & ARMA & WWR & ARMA benchmark      & ARMA benchmark &
    \Sexpr{myround(wwr_benchmark,tab_signif)} 
  \\
 \hline
\end{tabular}
\end{table}

>>>>

\section{Results}
\begin{table}[ht] % The placement specifier can be [h!tbp]
\centering 
\caption{AIC for the OvOv rotavirus model, computed using two filtering methods. PAL is the Poisson approximate likelihood, implemented using the code of WWR. PF is the particle filter, implemented using the R package pomp \citep{pomppackagepaper}. Lines 1, 2 and 7 are from WWR, and the remainder are our own computations. We used $5 \times 10^4$ particles for both PF and PAL. PF was repeated 36 times to reduce the Monte Carlo variance, but this step was not necessary for PAL due to its lower Monte~Carlo variance. PF results were maximized using iterated filtering, following the approach of SBH. PAL results were maximized using coordinate gradient descent, using the code of WWR.
} 
\label{tab:ovovrealdata}
\begin{tabular}{lllllr} 
  \hline
  & Method & Data & Model & Parameters                          & AIC  \\
  \hline
  1. & PF  & SBH  & SBH   & Table~1 of WWR, originally from SBH &  \Sexpr{myround(20134.38,tab_signif)}
  \\
  2. & PAL & WWR  & WWR   & Table~1 of WWR                      & \Sexpr{myround(13778.08,tab_signif)}
  \\
  3. & PAL & WWR  & WWR   & Re-evaluating the data, parameters from WWR &
    \Sexpr{myround(-2*palfilter_lik_list_wwr$log_lik + 2*7 ,tab_signif)} 
  \\
  4. & PF  & WWR & WWR    & Parameters matching 3. &
    \Sexpr{myround(-2*logmeanexp(logLik(pfilter_wwr)) + 2*7,tab_signif)}
  \\
  5. & PF  & WWR & WWR modified   & Initial values changed to match SBH &
     \Sexpr{myround(-2*est_mle_vec_sbh["logLik"] + 2*7,tab_signif)}
  \\
  6. & PAL & WWR & WWR modified   & Data, model and parameters matching 5.  &
     \Sexpr{myround(-2*palfilter_mif2_mle$log_lik + 2*7 ,tab_signif)}
  \\   
  7. & ARMA & SBH & ARMA benchmark    & Sec.~7.3.2 of WWR & 23043
  \\
  8. & ARMA & WWR & ARMA benchmark      & ARMA benchmark &
    \Sexpr{myround(wwr_benchmark,tab_signif)} 
  \\
 \hline
\end{tabular}
\end{table}
